Deriving an Executable Model of a Data Stream Protocol from its Specification

### 2009-10-05 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
paper ModelBasedTestCaseGenerationDataStreamProtocol: error detection results added

%
% "The UCD CASL SenseTile System", for IEEE Pervasive Computing
% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz 
%

@@

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan
Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

@@

\newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:} #1}}
\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:} #1}}
\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:} #1}}

\newcommand{\todo}{\textbf{TODO:}}

\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STSB}{\ST \emph{Sensor Board}\xspace}
\newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
\newcommand{\STPU}{\ST \emph{Processor Unit}\xspace}
\newcommand{\STPUs}{\ST \emph{Processor Units}\xspace}
\newcommand{\STU}{\ST \emph{Unit}\xspace}
\newcommand{\STUs}{\ST \emph{Unit}\xspace}

@@

\author{Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org, vieri.delbianco@gmail.com, dragan.stosic@gmail.com\\
}

@@

\todo explain and refine the problem

\todo search: other approaches to similar problems

@@

Finally in section \ref{sec:conclusion} conclusions are drawed and 
ongoing and future works on similar topics is introduced.

@@

The protocol is an asynchronous fixed length packed based protocol.
The protocol has been divided in multiple layers to ease its formal description.
The rationale is the following: mixing up different abstraction levels of a 
protocol description can lead to overly complex and unmanageable descriptions, 
expecially when dealing with formal ones.
The layers identified in protocol are the following:

@@

The single packet rules delimit the boundaries of the values obtainable from 
the packet: each defined sensor represented in the packet, as well as the 
metadata describing the \STSB and the packet and frame contents,  have a defined 
range of acceptable values.
There are also rules affecting more than one value, basically validity rules (i.e. if 
a specific sensor is active than the reading value has to be in a defined range), 
and rules specifing a correct sequence of frames (i.e. there can be only one switch 
in sampling rate in the frame sequence of a single packet).

@@

We also needed a specification capable to tackle properly all the protocol layers.
Since the different layers have very different characteristics, different context
dependant specification approaches were chosen.

@@

The specification is provided both by the Java API that has been built, and by the 
annotation with JML language\footnote{
\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}}. 
JML has been chosen because it is capable to specify Java code: JML is a behavioral 
interface specification language. 
JML specification language combines the design by contract approach (applayed to 
Java language) and the model-based specification approach of interface specification 
languages.

\todo insert proper JML references

Many tools are available supporting JML specification language, focusing on different 
aspects of the language. 
JML2 suite of tools
\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}}
was chosen, since both run time verification support and test generation support are 
provided directly by the suite.

\todo insert proper JML2 references

@@

The capabilities of a JML specification have been already descibed in the previous 
paragraph, that is: specification checkable at run time, tests generation, simulator 
verification, usable with hand written tests.

@@

Theoretically a separate simulator should be needed for each protocol layer; a
separate simulator for each layer would enable us to separeately specify, develop and test
each layer of the protocol.
In our case we evaluated not worth the effort to maintain such a separation in the 
simulators, motivated by the fact that the protocol layers identified are too fine grained 
to be separated into different simulators. 
Two simulators were built.

The two simulators were built reflecting the software abstraction layers defined for the 
driver.
The driver API is splitted in two structurual layers: the general high level interfaces, that 
expose the main functionalities of the board and the main contents of the packets hiding the 
details, and the lower level implementation that parse the data streams in and out the board, 
to translate the higher level intructions and data in properly formed packets.
Hence, we have implemented two simulators: one high level simulator, whicj reimplements the 
high level interfaces providing corresponding high level mechanisms to control the data and 
the behaviour of the simulated board, it has nothing to do with the details of the data 
streams.
One low level simulator which is capable to rebuild the data stream out of the board based on 
the data stream into the board, the in and out data streams are built exaclty as the sensor 
board is expected to be able to parse or generate.

@@

The JML specifications used to annotate Java code are of varing complexities.
Some of the specifications are simple, focusing on constraints that should hold when 
calling a method (preconditions) and the constraints on the return value (a very basic form of 
postconditions).
In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the 
method \lil{getTemperature} is declared to be \lil{/*@ pure @*/}, which means that 
it cannot change the state of any object (a postcondition), the specification also constraints 
the return value with a lower and upper bound (another postcondition).

@@

Some of the specifications are rather complex, usually focusing on properties regarding the 
behaviour of a whole object.
In listing \ref{lst:jml_example_invariant} a complex invariant 
(a property that has to be maintained, that is, that is assumed on entry and guaranteed on 
exit of each method of an object)
example is shown. 
The invariant is constraining the number of samples for each medium data rate streams: 
the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is 
\lil{FRAMS}, the constant constraining the number of samples is 
\lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are
fairly distributed on the channels. 
The valid samples are counted parsing all the frames contained in a packet, selecting only 
the matching valid samples.
A medium data rate stream sample is considered valid when \lil{isADCActive()} method 
returns \lil{true}.

@@

The unit test cases that have been built to verify the protocol driver 
are of two kinds: hand made unit tests and unit tests automatically 
generated based on JML specifications.
The tests have been tought to be used together, complementarly. 
Despite of this premise, the tests effectiveness has been evaluated 
separately and together, the evaluation is carried out in section 
\ref{sec:test_cases_retrospectives}.
The test effectiveness evaluation has been based on three elements:
code coverage (quantitative), development help and usefulness 
(qualitative), effective results on piloting the real board 
(quantitative).

The dual testing approach was thought as a winning option when 
compared to hand based tests or completely automated tests, in testing 
a driver for a stream communication protocol with hard constraints 
on development time.
\ST stream communication protocol is characterized by a well definite 
packet structure, and a well definite semantic on the packet contents 
as well.
On the other hand there are not strong constraints on synchronization, 
since there exists two concurrent communication channels and the protocol 
is asynchronous by construction.
The lack of hard synchronization constraints reduce the advantage to use 
a classical state machine based specification.

\todo reference to protocol tests generated from a state machine based 
specification is needed

Hard development time constraints were also a precise requirement; the time 
constraints pushed us towards a specification \emph{just good enough} for
our purposes.
A \emph{just good enough} specification is simply a specification that is not 
specifing everything, but only the software parts that are considered most 
critical, or the ones that are really easy to specify with the chosen 
specification language.
In \STSB driver JML was used to specify very simple behaviours (method 
oriented) on the whole driver implementaion, and more complex behaviour on 
the driver core interfaces.
The not covered areas were addressed by hand made unit tests.

Hand made unit tests cover behaviour not covered by the automatically 
generated tests, that is, they cover the not specified behaviour; but this 
is not the only purpose they have.
Hand made unit tests are used to drive the development, using the well known 
Test Driven Development approach, as well as to drive the most error prone 
specification as well.
Tests are used to drive the specification, a hand made unit test that should 
fail if the specification is not working as expected.

The rationale is the following: a stream communication protocol is 
not completely focused on a synchronised interaction

\begin{itemize}
 \item 
\end{itemize}

\todo describe dual approach: hand made tests + generated from 
specification tests

\todo reason about the usual structure oracle-driver-test-stub(s). a simulator is sometimes a 
stub, sometimes a driver, sometimes an oracle...

\todo insert references to testing embedded systems with simulators and/or with driver/stubs

\todo describe hybrid agile and TDD approach: high level informal 
specifications with BON, iterative refinements of code, to distill 
test cases and formal specification of the interfaces

\notev{maybe this part is to be saved for a different paper? BON has 
not been used properly}

\todo describe the necessity to add specification generated tests: to ensure
that specification is sound, to obtain higher coverage

@@

\todo describe hand made tests structure

\todo describe abstract test case and abstract test suite patterns: how to 
write tests against interfaces and/or abstract classes

@@

\todo describe the tool that generates the testing

\todo describe the structure of the generates tests

@@

\todo report high level analysis of the test suites

\todo describe measures: coverage, number of new bugs found

\todo compare hand made vs generated tests

\todo compare compare coupled approach vs single approaches

@@

\todo describe quantitative: coverage

\todo describe qualitative: errors found, help in design, ...

@@

\todo benefit: the approach can be used to test different part of the system
that share the same specifications

\todo testing the protocols layers

\todo testing the simulators

\todo describe general approach: depending at what level (of abstraction) tests 
and specifications are written, it is possible to verify both the simulator and 
the real protocol (isn't this obvious?)

@@

\todo

>>>>

%
% "The UCD CASL SenseTile System", for IEEE Pervasive Computing Joseph
% R. Kiniry, Vieri del Bianco, Dragan Stosic $Id: paper.tex 2795
% 2007-09-23 19:35:53Z dmz
%

@@

\ifpdf
\usepackage{epstopdf}
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Vieri del Bianco"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

@@

%\newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:} #1}}
\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:} #1}}
\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:} #1}}
\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:} #1}}

\newcommand{\todo}[1]{\texttt{\textbf{TODO:} #1}}

\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\SB}{\emph{Sensor Board}\xspace}
\newcommand{\STSB}{\ST \SB\xspace}
\newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
\newcommand{\STPU}{\ST \emph{Processor Unit}\xspace}
\newcommand{\STPUs}{\ST \emph{Processor Units}\xspace}
\newcommand{\STU}{\ST \emph{Unit}\xspace}
\newcommand{\STUs}{\ST \emph{Unit}\xspace}

@@

\author{Joseph R. Kiniry, Vieri del Bianco\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org, vieri.delbianco@gmail.com\\
}

@@

\todo{explain and refine the problem}

\todo{references: search other approaches to similar problems}

@@

Finally in section \ref{sec:conclusion} conclusions are drawn and 
ongoing and future works on similar topics is introduced.

@@

The protocol is an asynchronous fixed length packed based protocol.
The protocol has been divided in multiple layers to ease its formal description.
The rationale is the following: mixing up different abstraction levels of a 
protocol description can lead to overly complex and unmanageable descriptions, 
especially when dealing with formal ones.
The layers identified in protocol are the following:

@@

The single packet rules delimit the boundaries of the values obtainable from 
the packet: each defined sensor represented in the packet, as well as the 
metadata describing the \STSB and the packet and frame contents,  have a defined 
range of acceptable values.
There are also rules affecting more than one value, basically validity rules (i.e. if 
a specific sensor is active than the reading value has to be in a defined range), 
and rules specifying a correct sequence of frames (i.e. there can be only one switch 
in sampling rate in the frame sequence of a single packet).

@@

We also needed a specification capable to tackle properly all the protocol layers.
Since the different layers have very different characteristics, different context
dependent specification approaches were chosen.

@@

The specification is provided both by the Java API that has been built, and by the 
annotation with JML language\cite{JML2}\footnote{
\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}}. 
JML has been chosen because it is capable to specify Java code: JML is a behavioral 
interface specification language. 
JML specification language combines the design by contract approach (applied to 
Java language) and the model-based specification approach of interface specification 
languages.

Many tools are available supporting JML specification language, focusing on different 
aspects of the language. 
JML2 suite of tools\cite{BurdyEtal05-STTT}
\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}}
was chosen, since both run time verification support and test generation support are 
provided directly by the suite.

@@

The capabilities of a JML specification have been already described in the previous 
paragraph, that is: specification checkable at run time, tests generation, simulator 
verification, usable with hand written tests.

@@

Theoretically a separate simulator should be needed for each protocol layer; a
separate simulator for each layer would enable us to separately specify, develop and test
each layer of the protocol.
In our case we evaluated not worth the effort to maintain such a separation in the 
simulators, motivated by the fact that the protocol layers identified are too fine grained 
to be separated into different simulators. 
Two simulators were built.

The two simulators were built reflecting the software abstraction layers defined for the 
driver.
The driver API is splitted in two structural layers: the general high level interfaces, that 
expose the main functionalities of the board and the main contents of the packets hiding the 
details, and the lower level implementation that parse the data streams in and out the board, 
to translate the higher level instructions and data in properly formed packets.
Hence, we have implemented two simulators: one high level simulator, which re-implements the 
high level interfaces providing corresponding high level mechanisms to control the data and 
the behavior of the simulated board, it has nothing to do with the details of the data 
streams.
One low level simulator which is capable to rebuild the data stream out of the board based on 
the data stream into the board, the in and out data streams are built exactly as the sensor 
board is expected to be able to parse or generate.

@@

The JML specifications used to annotate Java code are of varying complexities.
Some of the specifications are simple, focusing on constraints that should hold when 
calling a method (preconditions) and the constraints on the return value (a very basic form of 
postconditions).
In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the 
method \lil{getTemperature} is declared to be \lil{/*@ pure @*/}, which means that 
it cannot change the state of any object (a postcondition), the specification also constraints 
the return value with a lower and upper bound (another postcondition).

@@

Some of the specifications are rather complex, usually focusing on properties regarding the 
behavior of a whole object.
In listing \ref{lst:jml_example_invariant} a complex invariant 
(a property that has to be maintained, that is, that is assumed on entry and guaranteed on 
exit of each method of an object)
example is shown. 
The invariant is constraining the number of samples for each medium data rate streams: 
the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is 
\lil{FRAMES}, the constant constraining the number of samples is 
\lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are
fairly distributed on the channels. 
The valid samples are counted parsing all the frames contained in a packet, selecting only 
the matching valid samples.
A medium data rate stream sample is considered valid when \lil{isADCActive()} method 
returns \lil{true}.

@@

The unit test cases that have been built to verify the protocol driver
are of two kinds: hand made unit tests and unit tests automatically
generated based on JML specifications.  The tests have been thought to
be used together, they are complementary.  The tests effectiveness has
been evaluated, for each kind of test, the evaluation is carried out
in section \ref{sec:test_cases_retrospectives}.  The test
effectiveness evaluation has been based on three elements: effective
results on piloting the real board (quantitative), code coverage
(quantitative), development help and usefulness (qualitative).

The dual testing approach was thought as a winning option, when 
compared to hand based tests or completely automated tests used alone, 
testing a driver for a stream communication protocol with hard constraints 
on development time.
\ST stream communication protocol is characterized by a well definite 
packet structure, and a well definite semantic on the packet contents 
as well.
On the other hand in the protocol there are not strong constraints on 
synchronization, since the communication channel is splitted in two (in and 
out) separate concurrent and asynchronous channels.
The lack of hard synchronization constraints reduce the advantage to use 
a classical state machine based specification \todo{references: 
protocol tests generated from a state machine based specification}.

Hard development time constraints were also a precise requirement; the time 
constraints pushed us towards a incomplete (\emph{just good enough}) 
specification.
In \STSB driver JML was used to specify very simple behaviors (method 
oriented) on the whole driver implementation, and more complex behavior on 
the driver core interfaces.
The software system parts not specified by the JML2 specification were 
specified using hand made unit tests.

Hand made unit tests cover behavior not covered by the automatically 
generated tests, that is, they cover the not specified behavior; but this 
is not the only purpose they have.
Hand made unit tests are used to drive the development, since the Test Driven 
Development approach\cite{beck2003test} has been followed during development.
Hand made tests are also used to drive the development of the most error 
prone or complex JML2 specifications; a test developed to drive the 
specification is a test that should fail if the specification is not working 
as expected.

\todo{describe the necessity to add specification generated tests: to ensure
that specification is sound, to obtain higher coverage}

@@

The package structure of the driver developed is shown in Figure 
\ref{fig:class_diagram_main}: two independent packages has been defined: 
\lil{Stream} and \lil{Driver}.
The packages are abstract, that is, they mainly contain abstractions; in Java 
language this is translated into a package which contains mainly abstract classes 
or interfaces.
The two abstract packages has various implementations, the dependency between
abstract packages is not direct, but only through an implementation 
(\lil{StreamDriver}).
This is needed to maintain a high decoupling of the packages, and is the result of 
applying the \emph{classical} dependency inversion 
principle\cite{martin1996dependency} (the dependency structure of the 
packages has a saw shape, a common shape on systems where the dependency 
inversion principle has been followed).

\begin{figure}[htb!]
 \centering
 \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__Main}
 \label{fig:class_diagram_main}
 \caption{Main packages class diagram.}
\end{figure}

The test package structure, which is shown in Figure 
\ref{fig:class_diagram_test} reflects and mimics the code structure; the
tests that test an abstract package are abstract, to be implemented by the 
package that tests a corresponding system implementation; this is a well 
known test pattern, the Abstract Test Pattern\cite{thomas2004java}, it is 
used to test that the contracts are respected in the implementations.
For instance, package \lil{DriverT} contains abstract tests for the 
abstractions of package \lil{Driver}, package \lil{StreamDriverT} 
inherits the abstractions contained in \lil{DriverT} and makes them concrete, 
to test the corresponding implementation \lil{StreamDriver}; package 
\lil{StreamDriverT} also contains stand alone tests written specifically for
the implementation contained in \lil{StreamDriver}.

\begin{figure}[htb!]
 \centering
 \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__Test}
 \label{fig:class_diagram_test}
 \caption{Test packages class diagram.}
\end{figure}

@@

The JML specification has been used to generate the tests. 
A specific tool of the JML2 suite can be used to automatically generate 
them\cite{Cheon-Leavens02}.

The test suite generated is a collection of fine grained unit tests.
For each public method a set of tests is generated.
The specified behavior of the method is checked using the JML2 run time 
assertion checker, hence, an oracle is not needed. 
The generated tests send messages to the method under test, they catch 
assertion violation exceptions from test cases that pass an initial 
precondition check; that is, if a precondition is not met, the test is 
not run, if the precondition is met the test is run, and if the postcondition 
is not me an exception is fired and the test fails.
The test data that is used to build the tests is the only part that is 
required to be provided by the user.

The test package structure is shown in Figure 
\ref{fig:class_diagram_generatedtest}.
The packages dependencies are defined by the specifications; for instance, 
\lil{SimulatorDriverST} are generated tests, they test (hence use) the 
implementation \lil{SimulatorDriver}, they have been generated using the 
specification corresponding specification \lil{SimulatorDriverS}.
The generated tests do not maintain the inheritance structure of the code 
and the specification.

\begin{figure}[htb!]
 \centering
 \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__GeneratedTests}
 \label{fig:class_diagram_generatedtest}
 \caption{Generated test packages class diagram.}
\end{figure}

\paragraph*{On simulators, again}

To have proper unit test suite, for each method under test a proper stub, 
driver and oracle should be identified, these are the usual components found 
in the harness that enable test execution\cite{binder1999testing}.
In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, the framework used 
to manage the implementation and execution of hand made and generated unit 
tests, there is no clear distinction between drivers, stubs, and oracle of a 
test; noneless, even if a proper distinction is not imposed by the framework, 
the conceptual distinction still holds.

Hence, to properly unit test a module in isolation and automatically we need:

\begin{description}
 \item[Stub or Skeleton]: a piece of code that simulates the activity of 
the modules that are not under tests.
 \item[Driver]: a piece of code that calls in a proper way the module under 
test.
 \item[Oracle]: a mechanism used for determining whether a test has passed 
or failed, comparing the output of the module under test to the output that 
the oracle determines that the module should have.
\end{description}

Hand made unit tests and JML generated tests differs in how these components 
are provided. 
The distinctions and similarities are summarized in Table 
\ref{tab:test_harness}.

\begin{table}[htbp]
\caption{Hand made and generated tests harness.}
\label{tab:test_harness}
\begin{center}
\begin{tabular}{|l|l|l|}\hline
& \textbf{\textit{hand made}} & \textbf{\textit{JML}} \\\hline
\textbf{\textit{Stub}} & to be provided & to be provided\\\hline
\textbf{\textit{Driver}} & to be provided & automatic: 1 call\\\hline
\textbf{\textit{Oracle}} & to be provided & 
automatic: runtime assertion checker\\\hline
\end{tabular}
\end{center}
\end{table}

A module simulator can be used to ease the tasks of developing each component 
of a test harness; that is, it can be used as a stub, to simulate a module 
needed by the module under test, it can be used as a driver, to stimulate the 
module under test, it can be used as an oracle, if it simulates the module 
under test.

In our system, the simulators has been used mainly as stubs, so that we have,
during testing, a structure like the one shown in Figure 
\ref{fig:class_diagram_streamdriver_test}.
In Figure \ref{fig:class_diagram_streamdriver_test} the test suite 
\lil{StreamDriverT} is using the \lil{SimulatorStream} to properly simulate 
the \lil{Stream} that is used by the system under test, that is 
\lil{StreamDriver}.

\begin{figure}[htb!]
 \centering
 \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__StreamDriverT}
 \label{fig:class_diagram_streamdriver_test}
 \caption{StreamDriver test packages class diagram: use of the stream simulator.}
\end{figure}

In our case, the usage of simulators to enable unit testing, was essential to 
test the \STSB driver, since the physical board was not available.
The simulator were used in both the hand made and the generated tests, since
both of them need a proper set of stubs to be executed.

\todo{insert references to testing embedded systems with simulators and/or 
with driver/stubs}

\paragraph*{BON}

\todo{describe approach: high level informal specifications with BON, 
iterative refinements of code, to distill test cases and formal 
specification of the interfaces} \notev{remove, BON has not been used 
properly, maybe a follow up paper}

@@

The use of both test suites, hand made and generated, has proven 
effective on finding protocols errors on the \STSB when the board 
prototypes have been delivered.

The first \STSB prototype delivered was meant to build empty packets, 
but the packets were built in a way not conformant to the 
specification, the sentinel values used to delimit a packet were not 
correct.
The board implementation error has been detected by the 
\lil{StreamDriver} package.

The second prototype delivered was meant to build packets with values
taken from rear sensors installed on the board.  The only sensors that
were not yet working were the sensors devoted to the fast rate data
streams, that is, the audio sensors.  Only one error needed to be
corrected in the driver implementation, because of a (rare)
combinations of conditions that were not covered by the test cases.
One error have been discovered in the low level driver that the driver
was using; to eliminate the error an upgrade of the low level drover
to a new unreleased version was necessary.  No other errors were
detected on the driver side. All the errors on the \SB side have been
detected.

The formal model and the run time assertion checker enabled the
detection of errors on the \SB itself.  Errors regarding the expected
values and values ranges of the filds contained in the packets and the
frames (4 protocol errors), and errors regarding the dependencies
between values in field (2 protocol errors) were detected. No other
errors, other than the ones immediately discovered by the run time
assertion checker were later on discovered.

The hand made test suite is composed by over 140 tests, while the
generated one is 100 times bigger, totalling over 14000 tests.

\todo{describe why such a difference: random selection, many tests 
pruned if pre condition is not satisfied}

\todo{compare hand made, many calls, vs generated, single call, tests}

\todo{describe measures: coverage, number of new bugs found}

\todo{compare compare coupled approach vs single approaches}

@@

\todo{describe quantitative: coverage}

\todo{describe qualitative: errors found, help in design, ...}

@@

\todo{benefit: the approach can be used to test different part of the system
that share the same specifications}

\todo{testing the protocols layers}

\todo{testing the simulators}

\todo{describe general approach: depending at what level (of abstraction) tests 
and specifications are written, it is possible to verify both the simulator and 
the real protocol (isn't this obvious?)}

@@
